---
title: "Data Analysis"
format: html
---

# Data Analysis Replication - Regional household variation and inequality across the Maya landscape

## Goal of the Original Paper

The main goal of Schroder et al. 2023 is to assess regional household variation and inequality across the Maya lowlands using a landscape-scale analysis. The authors apply Gini coefficients as a proxy to measure architectural differentiation—primarily household size—as an indicator of wealth inequality. Their study emphasizes the value and limitations of using lidar-derived data to interpret socio-economic patterns beyond elite centers.

## Data Set Used 

-   Lidar Data Source: NASA G-LiHT (Goddard's LiDAR, Hyperspectral, and Thermal Imager), collected over southeastern Mexico in 2013.

-   Coverage: 1,118 km² across multiple flight transects (\~300m wide, totaling 3,200 km in length).

-   Data Points: Annotated archaeological features (structures and patio groups) from 458 tiles.

-   Variables Used: Structure footprint area, structure volume, patio group area/volume, environmental and cultural factors (e.g., rainfall, site proximity, terrace density).

### Analyses Conducted 

The authors performed:

1.  Descriptive Statistics (e.g., structure size distributions, density, Gini coefficient ranges).

2.  Visualizations including:

-   Lorenz curves.

-   Correlation matrices.

-   Maps and density plots of structure and inequality distributions.

3.  Inferential Analyses:

-   Correlation testing between Gini coefficients and environmental variables (e.g., rainfall, proximity to sites).

-   Evaluation of regional and physiographic variation in inequality levels.

### Planned Replications

Figure 4: Descriptive statistics – assessing how sample size affects the Gini coefficient and its confidence intervals.

Figure 5: Visualization – map showing Gini coefficient distributions by tile.

Figure 9: Inferential statistics – correlation matrix between Gini coefficients and environmental variables (e.g., distance to water, site proximity, rainfall).

## Load Libraries

```{r}
library(tidyverse)
#install.packages("DescTools")
library(DescTools)
library(ggplot2)
#install.packages("ggcorrplot")
library(ggcorrplot)
#install.packages("sf")
library(sf)
#install.packages("ggmap")
library(ggmap)
library(readr)
library(dplyr)
```

Load data

Supplementary Data 1: Results of Gini analyses by G-LiHT tile
Supplementary Data 2: Correlation coefficients for Figure 8
Supplementary Data 3: P-values for Figure 8
Supplementary Data 4: Correlation coefficients for Figure 9
Supplementary Data 5: P-values for Figure 9
Supplementary Data 6: Correlation coefficients and p-values for Figures 10 and 11

```{r}
mmc1 <- read_csv("1-s2.0-S0278416523000685-mmc1.csv", col_names = TRUE)
mmc2 <- read_csv("1-s2.0-S0278416523000685-mmc2.csv", col_names = TRUE)
mmc3 <- read_csv("1-s2.0-S0278416523000685-mmc3.csv", col_names = TRUE)
mmc4 <- read_csv("1-s2.0-S0278416523000685-mmc4.csv", col_names = TRUE)
mmc5 <- read_csv("1-s2.0-S0278416523000685-mmc5.csv", col_names = TRUE)
mmc6 <- read_csv("1-s2.0-S0278416523000685-mmc6.csv", col_names = TRUE)
```
# Replication

## Figure 4
Fig. 4. Effect of sample size on a) difference between corrected and uncorrected
Gini coefficients based on structure footprint area and b) upper and
lower confidence intervals of Gini coefficients based on structure footprint area.

To replicate figure 4 we need supplementary material 1 (mmc1)

```{r}
head(mmc1) #view first few rows of data
```
Prepare the data

```{r}
# Rename columns for clarity
#df$Sample_Size <- df$`Sample Size Structure.x`
df$Gini_Uncorrected <- df$`Gini Structure Area`
df$Gini_Corrected <- df$`Corrected Gini Structure Area`

# Calculate Gini difference
df$Gini_Diff <- df$Gini_Corrected - df$Gini_Uncorrected
```
